Deciding Bitvector Termination with Program Synthesis

### 2014-07-05 papers/termination/paper.tex

\noindent {\bf Under-approximating the environment for non-termination.}
Dual to termination, when attempting to prove non-termination, we must under-approximate the environment.

@@

\section{Experiments}
We implemented our method by backing onto the {\sc Kalashnikov} program synthesiser.
We ran the tool on 40 benchmarks taken from the termination category of SVCOMP'15.
The experiments were performed on a 4-core 3.3GHz i5-2500k with 8GB of RAM.

>>>>

\noindent {\bf Under-approximating the environment for non-termination.}
Dual to termination, when attempting to prove non-termination, we must under-approximate the environment.
Consider the example in Figure~\ref{fig:nonterm-environment}.  We are attempting to prove non-termination
of the innermost loop, which we do by finding a recurrence set $N_4$.  We will assume that the other two
loops iterate only a finite number of times.  If the constraint system is satisfiable, $x_0$ is some initial
state in which the program begins executing.  Then $N_1$ is an inductive invariant of $L_1$ that is strong
enough to establish $N_2$ if $L_1$ exits (which it does by assumption).  Now, $N_2$ is an inductive invariant
of $L_2$ that is strong enough to establish either $N_3$ or $N_4$ on entry to $L_3$.  Recall that $N_4$ is
a recurrence set for $L_3$, so if a state $x \in N_4$ is reached, the loop $L_3$ will execute forever.
Otherwise, $N_3$ holds on entry to $L_3$.  $N_3$ is an inductive invariant of $L_3$ that is strong enough to
re-establish $N_2$ when $L_3$ exits.  Since we assumed that $L_2$ iterated only a finite number of times,
$N_3$ must be established only a finite number of times, which means that $N_4$ will eventually be reached,
at which point $L_3$ will loop forever since $N_4$ is a recurrence set for $L_3$.

\begin{figure*}
\begin{framed}
 \begin{minipage}{0.19\textwidth}
\begin{lstlisting}
L1: while (c1) {
     B1;
    }

L2: while (c2) {
     P;

L3:  while (c3) {
       B3;
     }
    }
\end{lstlisting}
\end{minipage}
\vline
\begin{minipage}{0.82\textwidth}
\begin{align*}
 \exists N_1, N_2, N_3, N_4, x_0 . \forall x, x' . \\
  N_1(x_0) \, \wedge \\
  c_1(x) \wedge N_1(x) \wedge B_1(x, x') & \rightarrow N_1(x') \, \wedge \\
  \lnot c_1(x) \wedge N_1(x) & \rightarrow N_2(x) \, \wedge \\
  N_2(x) & \rightarrow c_2(x) \, \wedge \\
  c_2(x) \wedge N_2(x) \wedge P(x, x') & \rightarrow N_3(x') \vee N_4(x') \, \wedge \\
  c_3(x) \wedge N_3(x) \wedge B_3(x, x') & \rightarrow N_3(x') \, \wedge \\
  \lnot c_3(x) \wedge N_3(x) & \rightarrow N_2(x) \, \wedge \\
  N_4(x) & \rightarrow c_3(x) \, \wedge \\
  N_4(x) \wedge B_3(x, x') & \rightarrow N_4(x')
\end{align*}
\end{minipage}
\end{framed}

\caption{The constraint system for non-termination\label{fig:nonterm-environment}}
\end{figure*}

@@

\section{Experiments}
We implemented our algorithm in the {\sc Headshot} tool, which uses the {\sc Kalashnikov} program synthesiser for termination.
We ran the tool on 40 benchmarks taken from the termination category of SVCOMP'15.
The experiments were performed on a 4-core 3.3GHz i5-2500k with 8GB of RAM.
